Analysing intention of chatbots use after COVID pandemic using Machine Learning

### 2021-06-27 edubots_surveys_future_paper.Rnw
created file for future paper

\title{Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective}

@@

\begin{abstract} % abstract
Introducing new technologies such as messaging platforms, and chatbots
attached to it, in higher education, requires careful consideration of
all options, but above all, an examination of the attitudes of the
affected community, especially those that will need to handle them,
% Antonio - I would say 'to manage them'. Students should also handle new technologies. ;)
% We're talking about teachers here, who need to handle the outcomes and introduce them in class - JJ
teachers.
In this paper we have surveyed the opinions of tertiary education
teachers based mainly in Spain and Spanish-speaking countries, 
looking at what they think about the introduction 
% Antonio - inthe classroom
% that's implied - JJ
of messaging platforms and chatbots, what their needs are, what kind of use cases they see for
them, and in what kind of environment they think its use will be
valuable. We will examine how and when these attitudes vary across
gender, experience, and other factors, such as the broad discipline
they are teaching. Our conclusions might help with the adoption of
messaging platforms in an educational setting, as well as chatbots or
other kind of bots, by higher education institution, and how these
must be introduced to teachers so that it helps them achieve desired
educational outcomes. 
\end{abstract}

@@

\begin{keyword}
\kwd{Chatbots}
\kwd{Messaging platforms}
\kwd{Tutorship}
\kwd{Educational bots}
\kwd{Higher education}
% Antonio - add something related to "survey", "teachers' opinion" or so?
\end{keyword}

@@

% ********************** METHODOLOGY ****************************

\section{Methodology}
\label{sec:meth}

Our intention from the beginning was to try and gather a sufficient amount of responses in all sectors, as well as cover different sectors of higher education. This is why we created a Google Form, in Spanish, which was intended to be answered in a very short amount of time. Initially, our intention was to mirror the questions used in \cite{MoraChatbots2021}; however, an initial exploration of results led us to create a second survey.
% Questions should probably be included here.

The initial survey was done in two different forms, one for university and other for tertiary non-university teachers. Questions and responses were the same, except for the type of tertiary degrees that were taught. We got around 300 responses to this one, with two thirds of them coming from university teachers, the rest from high school teachers. In both cases, we used mainly mailing lists, as well as Telegram groups, to spread the link to the form. The university form was spread in Spanish (mainly Andaluc\'{\i}a and Galicia) universities, as well as Costa Rica and Mexico. The tertiary non-university teachers are mainly in Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of 2021, post-pandemic and while, at least in Spain, many universities had mandatory virtual teaching.

The second survey followed the same paths, only in this case there was a single questionnaire. It was filled by teachers that were also students in a teacher formation course (on the use of new technologies in higher education, around 1/4 of them) as well as University of Granada teachers to knew about it by emails from the authors, as well as from the vicedeanship for International Relations that included it in its newsletter. This means that there might be a higher proportion of 1) teachers with few years of experience 2) teachers from the computer science faculty and 3) teachers from Granada. We think, however that there's no explicit bias in this selection, although of course specific percentages will vary.

% ********************** RESULTS ****************************

\section{Results}
\label{sec:res}

% Describe the survey and population
We collected responses from 282 teachers from Spain and Spanish Speaking countries for the first survey (\textit{survey1} in Fig. \ref{figure:teachers_population}) and 205 for the second one (\textit{survey2}). The sample includes mostly teachers from university, although about 32\% and 8\% are from vocational education in the first and second survey respectively. Regarding gender, in total 61\% are male teachers and 32\% female (approximately 5\% chose not to disclose it). Finally, the responses are equally distributed attending to the teaching experience, showing 24\% of responses from teachers with 5 or less years of experience, 26\% for 6-15 years of experience, 30\% for 16-25, and 20\% for teachers with more than 25 years of experience.

% [pedro]  the way these (5) figures are shown is a bit strange: different sizes of plots...

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.3\textwidth}
        \centering
<<gender, echo=FALSE, fig.width=3.2, fig.height=2.5>>=
#ggplot(freq.Gender, aes(x=Gender, y=freqGender, fill=Gender)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Gender$source<-rep("survey1")
freqInteraction.Gender$source<-rep("survey2")
freqbinded.Gender <- rbind(freq.Gender, freqInteraction.Gender)
freqbinded.Gender$source <- factor(freqbinded.Gender$source) 
ggplot(freqbinded.Gender, aes(x = Gender, y=freqGender, fill = source)) + geom_bar(stat="identity", position = "dodge", show.legend = FALSE)+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@

%       \includegraphics[width=\textwidth]{figures/gender.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \centering
<<age,echo=FALSE, fig.width=3.2, fig.height=2.5>>=
#ggplot(freq.Age, aes(x=Age, y=freqAge, fill=Age)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Age$source<-rep("survey1")
freqInteraction.Age$source<-rep("survey2")
freqbinded.Age <- rbind(freq.Age, freqInteraction.Age)
freqbinded.Age$source <- factor(freqbinded.Age$source) 
ggplot(freqbinded.Age, aes(x = Age, y=freqAge, fill = source)) + geom_bar(stat="identity", position = "dodge", show.legend = FALSE)+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@

%       \includegraphics[width=\textwidth]{figures/age.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \centering
<<sector,echo=FALSE, fig.width=3.2, fig.height=2.5>>=
#ggplot(freq.Sector, aes(x=Sector, y=freqSector, fill=Sector)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Sector$source<-rep("survey1")
freqInteraction.Sector$source<-rep("survey2")
freqbinded.Sector <- rbind(freq.Sector, freqInteraction.Sector)
freqbinded.Sector$source <- factor(freqbinded.Sector$source) 
ggplot(freqbinded.Sector, aes(x = Sector, y=freqSector, fill = source)) + geom_bar(stat="identity", position = "dodge")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector.pdf}
    \end{minipage}
    \vspace{0.5mm}
    \begin{minipage}[t]{0.38\textwidth}
        \centering
<<experience, echo=FALSE, fig.width=4.5, fig.height=3.475>>=
#ggplot(freq.Experience, aes(x=Experience, y=freqExperience, fill=Experience)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Experience$source<-rep("survey1")
freqInteraction.Experience$source<-rep("survey2")
freqbinded.Experience <- rbind(freq.Experience, freqInteraction.Experience)
freqbinded.Experience$source <- factor(freqbinded.Experience$source) 
ggplot(freqbinded.Experience, aes(x = Experience, y=freqExperience, fill = source)) + geom_bar(stat="identity", position = "dodge", show.legend = FALSE)+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@        
%       \includegraphics[width=\textwidth]{figures/experience.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.52\textwidth}
        \centering
<<discipline,echo=FALSE, fig.width=8.2, fig.height=4.45>>=
#ggplot(freq.Discipline, aes(x=Discipline, y=freqDiscipline, fill=Discipline)) + geom_bar(stat="identity")+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
freq.Discipline$source<-rep("survey1")
freqInteraction.Discipline$source<-rep("survey2")
freqbinded.Discipline <- rbind(freq.Discipline, freqInteraction.Discipline)
freqbinded.Discipline$source <- factor(freqbinded.Discipline$source) 
freqbinded.Discipline$Discipline <- factor(freqbinded.Discipline$Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other"))
ggplot(freqbinded.Discipline, aes(x = Discipline, y=freqDiscipline, fill = source)) + geom_bar(stat="identity", position = "dodge", show.legend = FALSE)+scale_y_continuous(labels=scales::percent) + theme(axis.title.y=element_blank())
@

%       \includegraphics[width=\textwidth]{figures/discipline.pdf}
    \end{minipage}
\end{center}
    \caption{Sample population description. The responses from two surveys were collected amount to 282 for the first survey (\textit{survey1} in red) and 205 for the second (\textit{survey2} in blue). Data are shown by gender, age, sector (university and vocational education), number of years of experience in education, and discipline.}
\label{figure:teachers_population}
\end{figure}

Figure \ref{figure:teachers_population} shows the distribution of age and experience are quite similar. %, and as a matter of fact,
Additionally, as a matter of fact, as Figure \ref{figure:age.vs.exp} shows, age and experience are correlated strongly, with experience being roughly 20-30 years less than age. 
This might indicate that a differential analysis for age and experience will not yield significant differences, except maybe in the 6-15 years experience group, which is spread over all ages. At any rate, experience seems to be the dominating factor here and we'll use experience only unless we detect some significant difference by age.

\begin{figure}[h!tbp]
\begin{center}
<<interaction,message=FALSE,echo=FALSE, fig.height=2.0, fig.width=4.0>>=
library(forcats)
survey.interaction %>%
    mutate(Age = fct_relevel(Age,"25-35", "35-45", "45-55", "> 55")) %>%
    mutate(Experience = fct_relevel(Experience, "0-05 years", "06-15 years", "16-25 years", "> 25 years" )) %>%
    group_by( Age, Experience ) %>% summarise( n = n() ) %>% mutate( freq = n/sum(n)) -> Age.vs.Experience

ggplot(Age.vs.Experience,aes(Age,Experience))+geom_point(aes(size=freq),colour="green")
@
\end{center}
\caption{Correlation between age and experience.}
\label{figure:age.vs.exp}
\end{figure}

\subsection{RQ1 - Are teachers already using messaging apps in their classes?}

%Why messaging apps are relevant here --> connect to the development of chatbots
Group messaging apps have become a powerful tool for keeping touch with students. Also, they are the baseline platform for the development of chatbots to help with the learning process.

According to the results summarized in Fig. \ref{figure:chatbots_use}, teachers mainly use the instant messaging app provided by their institutions (55\%),  and Whatsapp (44\%), the most popular messaging app in Spain. Finally, about 20\% also use Telegram. 
Although gender differences are not significant, female teachers answered they use instant messaging apps more than male teachers (about 10\% more). Also, teachers in vocational Education also use Whatsapp more than university teachers.

With respect to the distribution of the use of messaging apps per age, there are no significant differences for the Whatsapp and apps provided by their institutions. However, younger teachers use also Telegram with more than 25\%, a percentage that falls to about 10\% for teachers that are 55 or older. The third column compares the use of these apps considering the number of years of experience. One interesting result is that, contrarily to the idea of technology adaptation issues when considering older teachers, about 65\% of teachers with more than 25 years of experience use the platforms provided by their institutions while the percentage goes down to less than 50\% for teachers with 6-15 years of experience.

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.31\textwidth}
        \centering
<<gender_provided, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformProvided, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male','PNTS')))+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_provided.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.31\textwidth}
        \centering
<<experience_provided, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformProvided, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Experience)+ theme(axis.title.y=element_blank())
@       
%       \includegraphics[width=\textwidth]{figures/experience_provided.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.31\textwidth}
        \centering
<<sector_provided, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformProvided, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank())
@       
%       \includegraphics[width=\textwidth]{figures/sector_provided.pdf}
    \end{minipage}
    %----
    \vspace{0.2mm}
    \hspace{10mm}
    \begin{minipage}[t]{0.5\textwidth}
        \centering
<<discipline_provided, echo=FALSE, fig.width=18.6/2.54, fig.height=7.7/2.54>>=

ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformProvided, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other")))+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/discipline_provided.pdf}
    \end{minipage}
    \hspace{10mm}
    %----
    %----
    \vspace{0.5mm}
    \begin{minipage}[t]{0.31\textwidth}
        \centering
<<gender_whatsapp, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformWhatsApp, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male','PNTS')))+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_whatsapp.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.31\textwidth}
        \centering
<<experience_whatsapp, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformWhatsApp, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Experience)+ theme(axis.title.y=element_blank())
@       
%       \includegraphics[width=\textwidth]{figures/experience_whatsapp.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.31\textwidth}
        \centering
<<sector_whatsapp, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformWhatsApp, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank())
@       
%       \includegraphics[width=\textwidth]{figures/sector_whatsapp.pdf}
    \end{minipage}
    \vspace{0.2mm}
    \hspace{10mm}
    \begin{minipage}[t]{0.5\textwidth}
        \centering
<<discipline_whatsapp, echo=FALSE, fig.width=18.6/2.54, fig.height=7.7/2.54>>=
ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformWhatsApp, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other")))+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/discipline_whatsapp.pdf}
    \end{minipage}
    \hspace{10mm}
    %----
    %----
    \vspace{0.5mm}
    \begin{minipage}[t]{0.31\textwidth}
        \centering
<<gender_telegram, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformTelegram, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male','PNTS')))+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_telegram.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.31\textwidth}
        \centering
<<experience_telegram, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformTelegram, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Experience)+ theme(axis.title.y=element_blank())
@       
%       \includegraphics[width=\textwidth]{figures/experience_telegram.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.31\textwidth}
        \centering
<<sector_telegram, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformTelegram, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank())
@       
%       \includegraphics[width=\textwidth]{figures/sector_telegram.pdf}
    \end{minipage}
    \vspace{0.2mm}
    \hspace{10mm}
    \begin{minipage}[t]{0.5\textwidth}
        \centering
<<discipline_telegram, echo=FALSE, fig.width=18.6/2.54, fig.height=7.7/2.54>>=
ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformTelegram, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other")))+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/discipline_telegram.pdf}
    \end{minipage}
    \hspace{10mm}
\end{center}
    \caption{Use of chatbots in class: distributions per gender, age, years of experience in education, university or vocational education, and discipline. Teachers were specifically asked about two popular instant messaging apps (Whatsapp and Telegram), and the use of platforms already provided by their institutions different than these two ones.}
\label{figure:chatbots_use}
\end{figure}

Results in Fig. \ref{figure:chatbots_use} do not show differences in the use of instant messaging apps between teachers from University and vocational education. In general, most teachers prefer messaging apps provided by their own institutions. With respect to specific disciplines, Engineering and Technology teachers are more active in their use, although it is also relevant the amount of teachers from Humanities that answered they used these apps in their classes (around 60\% use the apps provided by their institutions).

%\begin{figure}[t]
%\begin{center}
%    \vspace{2mm}
%    \begin{minipage}[t]{0.30\textwidth}
%	    \centering
% 	    \includegraphics[width=\textwidth]{figures/sector_whatsapp.pdf}
%    \end{minipage}
%    \begin{minipage}[t]{0.50\textwidth}
%	    \centering
% 	    \includegraphics[width=\textwidth]{figures/discipline_whatsapp.pdf}
%    \end{minipage}
%    
%    \vspace{2mm}
%    \begin{minipage}[t]{0.30\textwidth}
%	    \centering
% 	    \includegraphics[width=\textwidth]{figures/sector_telegram.pdf}
%    \end{minipage}
%    \begin{minipage}[t]{0.50\textwidth}
%	    \centering
% 	    \includegraphics[width=\textwidth]{figures/discipline_telegram.pdf}
%    \end{minipage}
%\end{center}
%    \caption{Differences in the use of chatbots in class between vocational and university teaching and distributions by discipline.}
%\label{figure:chatbots_use_sector_discip}
%\end{figure}

Regarding the social factor of chat groups that teachers use in class, according to Fig. \ref{figure:chatbots_messagingorganization} the vast majority prefer small groups only with the students from the same course. These are more focused groups with specific goals and dedicated to the organization of the course and its tasks, and from the pedagogic point of view seem more appropriate to improve the learning process. Interestingly, about 30\% of teachers consider they should not be part of the chat group. Also, only university teachers find interesting a group with all the students and teachers in their own Faculty or School. The lack of teachers from vocational Education here may be the consequence of using specific language such as "Faculty". Moreover, the fact that many universities and schools already use these groups for administrative and social interaction (e.g. \cite{dibitonto2018chatbot}) might be the reason for the low percentage of teachers that selected this response.

\begin{figure}[h!tbp]
\centering
<<messaging_organization, echo=FALSE, fig.width=19/2.54, fig.height=5/2.54>>=
survey.pilot.MessagingOrganization <- within(survey.pilot.MessagingOrganization, Messaging.Organization <- factor(Messaging.Organization, levels=names(sort(table(Messaging.Organization), decreasing=TRUE))))
ggplot(survey.pilot.MessagingOrganization, aes( x = Messaging.Organization, y = ..count.., group=Sector, fill = Sector ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90))+coord_flip()+ theme(axis.title.y=element_blank())
@
%\includegraphics[width=\textwidth]{figures/messaging_organization.pdf}
\caption{Distribution of teachers' preferences for the chat groups with their students: from groups only with their students from a specific course to groups with greater social interaction with all students in their School or Faculty.}
\label{figure:chatbots_messagingorganization}
\end{figure}

Some of the questions in the survey were focused on the impact of the COVID19 pandemic between 2020 and 2021, in the teachers' attitudes towards the instant messaging apps. The 273 answers are summarized in Fig. \ref{figure:chatbots_use_postcovid}, showing that about 77\% of teachers already used these tools before the pandemic and kept using them during the pandemic lockdowns that forced students and educators to use remote education schemes. Moreover, approximately 15\% of them switched their messaging app for one that offered a safer interaction with their students. According to the responses, an additional 16\% started using messaging apps during the pandemic for the first time in their classes. Aggregating the answers, currently 91\% of the teachers are using messaging apps as learning tools.

%A few relevant facts from the collected data: teachers with 16-25 years of experience seem more worried about the security of messaging apps and teachers with up to 5 years of experience were more open to adopt messaging apps due to the pandemic; female teachers were more willing to adopt new platforms due to the pandemic, and even to start using safer alternatives; finally, university teachers that did not use them were more reluctant to start using them after the pandemic.

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<gender_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
survey.pilot.postCOVID <- within(survey.pilot.postCOVID, PostCOVID.changes <- factor(PostCOVID.changes, levels=names(sort(table(PostCOVID.changes), decreasing=TRUE))))
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Gender, fill = Gender ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_covid.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<sector_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Sector, fill = Sector ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector_covid.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<experience_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Experience, fill = Experience ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@       
%       \includegraphics[width=\textwidth]{figures/experience_covid.pdf}
    \end{minipage}
\end{center}
    \caption{Impact of COVID in the use of messaging apps in University and vocational education.}
\label{figure:chatbots_use_postcovid}
\end{figure}

The descriptive analysis in this case is done using a Chi-square test. For the different variables considered, some statistical significance was found for Gender, Sector, and academic Discipline. The values found with the test are: $p-value=0.0034$ for Gender, $p-value=0.0199$ for Sector, and $p-value=0.0211$ for Discipline.
The first column in Fig. \ref{figure:postcovid_chisq} shows the Pearson's residuals of the test for these three variables. Residuals show the difference between the observed and the expected values, with large residuals showing that variables are not truly independent. In this case, blue shows positive contribution and red negative contribution to the test, and saturation shows how large is this contribution in contrast to the expected value by chance. Tables on the right side show adjusted standardized residuals that according to \cite{agresti2013categorical}, if greater than +/-2 for cases with few cells, indicate lack of fit of $H_0$ (bolded in our table). Also, the table shows the percentage of contribution to the test of each cell.

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[c]{0.45\textwidth}
        \centering
<<gender_postcovid_chisq, echo=FALSE, fig.width=17/2.54, fig.height=10/2.54>>=
df <-select(survey.pilot.postCOVID, PostCOVID.changes, Gender, Sector, Experience, Discipline)
df[as.character(df$Gender)!="PNTS", ] -> df1
df1[as.character(df1$Gender)!="No", ] -> df1
df1[as.character(df1$Discipline)!="No", ] -> df1
df1[as.character(df1$Experience)!="No", ] -> df1
df1[as.character(df1$Sector)!="No", ] -> df1
survey.pilot.postCovidChisq<- droplevels(df1)

covidChanges.vs.Gender<-table(survey.pilot.postCovidChisq$Gender, survey.pilot.postCovidChisq$PostCOVID.changes)
colnames(covidChanges.vs.Gender)<- c("Did/do", "Didn't/do","safer", "Didn't/don't")
chisq<-chisq.test(covidChanges.vs.Gender)
corrplot(chisq$residuals, is.corr = FALSE)
@
    \end{minipage}
    %This tables comes from (TODO: generate them automatically from the data)
    %contrib <- 100*chisq$residuals^2/chisq$statistic
    %round(contrib, 2)
    %round(chisq$stdres, 3)
    \begin{minipage}[c]{0.45\textwidth}
        \centering
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{l|r|r|r|r|}
    \cline{2-5}
                                 & Did/Do & Didn't/Do & Safer  & Didn' t/Don't \\ \hline
    \multicolumn{1}{|l|}{Female} & \textbf{-2.850} & \textbf{2.806} & 1.832  & -1.011 \\ 
    \multicolumn{1}{|l|}{} & 15.15\% & \textbf{31.21\%} &  13.49\% & 4.40\% \\ \hline
    \multicolumn{1}{|l|}{Male}   & \textbf{2.850} & \textbf{-2.806} & -1.832 & 1.011 \\ 
    \multicolumn{1}{|l|}{} &  8.43\% &  \textbf{17.36\%} & 7.50\% &  2.45\% \\ \hline
    \end{tabular}%
    }
    \end{minipage}
    \begin{minipage}[c]{0.45\textwidth}
        \centering

>>>>

\title{Analysing intention of chatbots use after COVID pandemic using Machine Learning}

@@

\begin{abstract} % abstract

\end{abstract}

@@

\begin{keyword}
\kwd{Chatbots}
\kwd{Messaging platforms}
\kwd{Tutorship}
\kwd{Educational bots}
\kwd{Higher education}
\end{keyword}

@@

% ********************** SURVEY DESCRIPTION ****************************

\sectionData description}
\label{sec:data}

% ************************ DATA ANALYSIS ***********************

\section{Data Analysis using Machine Learning}

To provide a different perspective another analysis was carried out using models from the field of machine learning. The idea was to be able to identify which features measured in the survey could identify willingness to change. As chatbots are an emerging technology, it is quite probable that this new change will come sooner than later considering the improvements in Virtual and Augmented Reality \cite{}.

The output variable is set as covid-changes, whose values were grouped to three numerical classes:

\begin{itemize}
 \item 2 = Yes, now they use it or use it more or in a safer/different way 
 \item 1 = No change in use 
 \item 0 = No change in not using 
\end{itemize}

Regarding the rest of features, a similar preprocessing was applied to the original survey:

\begin{itemize}
\item Binary Yes/No variables like AgendaBot, ClassMaterialBot, FAQBot, GradesBot, MessagingPlatform..., OtherBot were mappped to values $\in {0,1}$ using 1 for Yes and 0 for No.

\item Gender variable was mapped to: 1 for Male, 2 for PNTS and 3 for Female. In this way we keep equidistance from PNTS to the other well identified genders.

\item Age and Experience were assigned valued $\in {1,2,3,4}$ where the higher number represents the older age or more years of experience.

\item Sector was coded as 1 for University and 2 for Vocational.

\item Messaging Organization was reinterpreted as number of approaches considered by the professors with the following values: 0 for answer of type: "I'd rather leave students self-organize and not be part of the chat" ; 1 for answers like: "A chat with my students in the same course" or "A chat with all students and teachers in the same semester/year". If the previous two were marked, or even another answer appeared, the sum of approaches was computed. By doing this, we consider a passive professor that does not want to get involved in the chat with a 0 and emphasize the active ones that give a thought on how they want to interact with students.

\item Degrees feature was also reinterpreted as number of degrees involved. By doing this, it is possible to see with a number the range width of the professor. From this point, it could be possible to study if a wider range professor presents or not more intention of using chatbots (independently of the degree level).

\item Finally, Discipline was grouped into the following categories: 0 for No  1 for Engineering and Technology 2 for Science and Biomedicine ; 3 for Social Science and 4 for Humanities. By doing this we keep a similarity criterion just by computing the distance between disciplines.

The reference mapping for the feature names in the experiments is in Table \ref{tab:feature_mapping}. 
\begin{table}
\begin{tabular}{|c|c|}
\hline
\# Variable & Name \\
\hline
0 & AgendaBot	\\
1 & ClassMaterialBot		\\
2 & Age		\\
3 & Experience		\\
4 & FAQBot		\\
5 & GradesBot		\\
6 & Gender		\\
7 & MessagingPlatform0ne		\\
8 & MessagingPlatformOther		\\
9 & MessagingPlatformProvided		\\
10 & MessagingPlatformSlack		\\
11 & MessagingPlatformTelegram		\\
12 & MessagingPlatformWhatsApp		\\
13 & OtherBot		\\
14 & Sector		\\
15 & Messaging Organization $\rightarrow$ Number of approaches considered		\\
16 & Degrees $\rightarrow$ Number of degrees involved		\\
17 & Discipline		\\
\hline
Output & PostCOVID changes	\\
\hline
\end{tabular}
\caption{Mapping of feature numbers to survey questions.}
\label{tab:feature_reference} 
\end{table}

\subsubsection{Using Mutual Information and Neural Networks}

The intuitive idea for Mutual Information (MI) is the amount of information about a variable that another variable, or group of variables, has. Therefore, setting as goal the capability of predicting the COVID changes, it is possible to detect the most relevant variables. Several approaches can be taken to select a subset of variables that contain enough information to reconstruct the output \cite{}. For this study, the estimation of the mutual information provided by the SKlearn package \cite{} and the algorithm mMRM \cite{} are considered.

As the evaluation of all possible combinations might be unfeasible from a computational point of view, the following heuristic was applied: rank the variables using the MI of that variable isolated, apply a threshold to discard the ones with smaller MI and then, make a exhaustive search for the combination with higher MI.

\begin{table}
\begin{tabular}{|c|c|c|c|c|}
\hline
\# Variable & Name & MI (considering one variable) & Max Relevance & mRMR\\
\hline
0 & AgendaBot	&  &  &\\
1 & ClassMaterialBot &	&  &	\\
2 & Age	&		&  &  \\
3 & Experience &			&  &  	\\
4 & FAQBot &			&  &  	\\
5 & GradesBot &			&  &  	\\
6 & Gender &			&  &  	\\
7 & MessagingPlatform0ne &			&  &  	\\
8 & MessagingPlatformOther &			&  &  	\\
9 & MessagingPlatformProvided	&			&  &  \\
10 & MessagingPlatformSlack	&			&  &  \\
11 & MessagingPlatformTelegram	&			&  &  \\
12 & MessagingPlatformWhatsApp	&			&  &  \\
13 & OtherBot	&			&  &  \\
14 & Sector	&			&  &  \\
15 & Messaging Organization $\rightarrow$ Number of approaches considered	&			&  &  \\
16 & Degrees $\rightarrow$ Number of degrees involved	&			&  &  \\
17 & Discipline	&		&  &  	\\
\hline
Output & \multicolumn{3}{c}{PostCOVID changes} & 	\\
\hline
\end{tabular}
\caption{Mapping of feature numbers to survey questions.}
\label{tab:feature_results}
\end{table}

\subsubsection{Using XGBoost}

Due to the lack of data, estimators based on $k$-Nearest Neighbours might be unreliable, thus, another approach can be used: decision trees build during their design process a relevance of the variables that, after being used to take the decision, decrease in a significant way the error. Therefore, the well known implementation of the XGBoost \cite{} was used to obtain the accuracy and the feature relevance.

The feature ranking results are shown in Figure \ref{fig:feature_xgboost}.
% \begin{figure}
% \includegraphics{figs/feature_xgboost.png}
% \label{fig:feature_xgboost}\caption{Ranking of features obtained after using XGBoost model.}
% \end{figure}
